How to compute the fusion product of MV cycles in type $A$

### 2021-05-18 draft.tex
almost fixed

%!TEX program = lualatex

\documentclass[draft]{article}
\usepackage{basic}
\usepackage{emoji}
% TODO: The matrix is lower triangular and the lower-triangular matrix is
% \setemojifont{Apple Color Emoji}
% Working title: 
\title{How to compute the fusion product of MV cycles in type A}
\author{Roger Bai, Anne Dranowski, Joel Kamnitzer} 
% \date{October 2020}
\date{\today}

@@

\begin{theorem} 
\label{th:OGrl}
There is an isomorphism
$$
    \overline{\OO}^{\lambda', \lambda''}_{0,\AA} \cap \TT_\mu \cong \overline\Gr^{\lambda', \lambda''}_{0,\AA} \cap \cW_\mu 
$$
given by $ (A,s) \mapsto (g(A)\CC[t]^m, s)$.
\end{theorem}
% \acom{Joel, you stopped writing $X_{0,\AA}$ in favour of $X_\AA$ --- should we do this everywhere?}
% \jcom{Oops, no I think that we should stick with $ X_{0, \AA}$.}
% TODO: comb the doc at the end for this notation 
\begin{proof}
Since we already have the isomorphism from \Cref{th:TmuWmu}, it suffices to show that for any $ A \in \TT_\mu$, 
$$ 
    (A,s) \in \overline{\OO}^{\lambda', \lambda''}_{0,\AA} \cap \TT_\mu \text{ if and only if } (g(A)\CC[t]^m, s) \in \overline\Gr^{\lambda', \lambda''}_{0,\AA} \cap \cW_\mu \,. 
$$
This follows immediately from \Cref{le:Grl1l2}.
\end{proof}
% 
\begin{theorem}
\label{th:OTGrW}
The isomorphism from \Cref{th:OGrl} 
% \acom{should ``Theorem'' etc.\ be capitalized when it appears in the middle of a sentence?} Yup.
restricts to an isomorphism
$$ 
    \overline{\OO}^{\lambda', \lambda''}_{0,\AA} \cap \UU^{\mu', \mu''}_{0,\AA} \cong \overline{\Gr}^{\lambda', \lambda''}_{0,\AA} \cap S^{\mu', \mu''}_{0,\AA}\,.
$$
\end{theorem}
% 
%\jcom{I was thinking about the proof of this result.  It is enough to prove that for $ A \in \OO^{\lambda', \lambda''}_s \cap \TT_\mu $, $g(A) \in S^{\mu', \mu''}_s $ if and only if $ A \in T^+$.  It is easy to show that if $ A \in \UU$, then $ g(A) \in S^{\mu', \mu''}_s $.  But the converse is not so obvious.  (I think Anne's paper/thesis might be incomplete on this point.)  I thought of two approaches: first to write down the lattice interpretation of $S^{\mu', \mu''}$ (similar to what is in my thesis in the Anderson-Kogan comparison section) or think about both $S^{\mu', \mu''}$ and $\UU$ as attracting sets for a $\CC^\times$ action.  Which do you prefer?}
% 
%\acom{Assuming Theorem 2 we think this follows by application of Roger's lemma. Also, I think my thesis is complete on this point.}
% 
%\jcom{I thought about this and I agree with you now.  I will write this up soon.} 
% 
%\acom{But I am also curious about the attracting set approach.}
% 
\begin{proof}
We could prove this by observing the both sides are the attracting locus of an appropriate $ \CC^\times$ action. However, we will give the following more algebraic proof.

@@

Restricting to the zero fibre and applying $\tau_0$ we recover \cite[a main result]{dthesis}, itself a corollary of the ordinary affine Grassmannian version of \Cref{th:OGrl}, \cite[Theorem~4.4.1(c)]{mirkovic2019comparison}. 
% 
\begin{corollary}
    \label{cor:mvy}
    $\overline{\OO}^\lambda \cap \TT_\mu\cap\n \cong\overline\Gr^\lambda\cap S^\mu_+$ where $\n\subset M_N(\CC)$ is the subalgebra of upper-triangular matrices. 
\end{corollary}
%The closures of the irreducible components of the righthand side are \textbf{MV cycles of type $\lambda$ and weight $\mu$}. By \cite{baumann2019mirkovic} the set of MV cycles of all possible types and weights can be stitched together to form a \textbf{perfect basis} of $\CC[N]$.

@@

We begin with the following definitions.
\begin{definition}
Let $ L \subset \OO^m$ be a point in $ \Gr_+$.  We define the relative dimension of $L $ by
$$
\rdim L := \dim_\CC \OO^m/L\,. 
$$
% 
If $ \gamma \subset \{1, \dots, m\}$, we define two lattices $ L^\gamma, L_\gamma $ in $ \OO^\gamma := \Sp_{\OO}(e_i : i \in \gamma) $ by
\begin{gather*}
%    \OO^\gamma := \Sp_{\OO}(e_i : i \in \gamma) \\
L^\gamma:= L \cap \OO^\gamma  \qquad L_\gamma := L / L^{\gamma^c} \subset \OO^m / \OO^{\gamma^c} = \OO^\gamma
\end{gather*}
where $ \gamma^c := \{1, \dots, m \} \setminus \gamma$.
\acom{I suggest replacing $\mathbb O$ by $R$, and $L/L^{\gamma^c}$ by $(L + R^{\gamma^c})/R^{\gamma^c}$ because that's what we're doing. In a lower triangular $g(A)$ columns ${i+1..m}$ cannot be used to eliminate entries in rows ${i+1..m}$ in the first $i$ columns over $\CC[t]$ or $\CC\xt$.}
\jcom{If you think that it sounds better with $ \OO $ replaced by $ R$, then I won't object to the change. Note that the first sentence won't quite make sense ``let $ L \subset \OO^m$ be a point in $ \Gr_+$'', we would have to write ``let $ L \subset R^m$ be a lattice''.}
\jcom{$ L /L^{\gamma^c} $ is the image of $ L $ in $ \OO^m / \OO^{\gamma^c}$ since the kernel of the map $ L \rightarrow \OO^m / \OO^{\gamma^c}$ is precisely $L^{\gamma^c} $, so I don't think that there is any problem here.  Again, if you think that it is better, I won't object to your suggested change.} 
\jcom{Your question made me wonder about the following: how do we know that $ L_\gamma$ is a lattice?  (or $ L^\gamma$ for that matter)}
\end{definition}
We will also need the analogous definitions when $ \OO$ is replaced by $ \CC[t]$ and $L \in \Grth_+$.

We record the following easy observations.
\begin{lemma} \label{le:rdimTrans} Let $ L \in \Gr_+$.
\begin{enumerate}
    \item For any $ 
    \gamma \subset \{1, \dots, m\} $, we have $$ \rdim L_\gamma + \rdim L^{\gamma^c} = \rdim L \,.$$
    \item For any $ \gamma_1 \subset \gamma_2 \subset \{1, \dots, m\}$, we have $$ (L_{\gamma_1})_{\gamma_2} = L_{\gamma_2}
$$
\end{enumerate}
\end{lemma}

For any $ L \in \Gr_+ $ and $\gamma \subset \{1, \dots, m\} $, we let $ D_\gamma(L) := \rdim L^\gamma = \rdim L - \rdim L_{\gamma^c}$. By \cite[Proposition 9.3]{kamnitzer2010mirkovic}, this coincides with the general definition of $ D_\gamma$ given in that paper.

Recall that $ S^\mu_- = N_-(\cK) t^\mu$. We also introduce the notation $S^\mu_+ := N(\cK)t^\mu$.

\jcom{Actually we will need this notation later when we talk about stable MV cycles.  Maybe we should change $ S^\mu $ to $S^\mu_-$.}
\acom{For it.}

\begin{lemma} \label{le:Srdim}
    Let $ L \in \Gr_+$.  The following are equivalent:
    \begin{enumerate}
        \item $ L \in S_+^\mu$
        \item $ \rdim L^{\{1, \dots, i\}} = \mu_1 + \dots + \mu_i$ for all $ i = 1, \dots, m $.
        \item $ \rdim L_{\{i+1, \dots, m\}} = \mu_{i+1} + \dots + \mu_m$ for $ i  = 1, \dots, m$.
    \end{enumerate}
    Similarly, the following are equivalent:
        \begin{enumerate}
        \item $ L \in S_-^\mu$
        \item $ \rdim L^{\{i+1, \dots, m\}} = \mu_{i+1} + \dots + \mu_m$ for all $ i = 1, \dots, m $.
        \item $ \rdim L_{\{1, \dots, i\}} = \mu_{1} + \dots + \mu_i$ for $ i  = 1, \dots, m$.
    \end{enumerate}
\end{lemma}

We now introduce some notation related to Young tableaux. Let $YT(\lambda)_\mu$ denote the set of (possibly semistandard) Young tableaux of shape $\lambda$ and weight $\mu$. 
%Here, the weight of a tableau is a sequence $(\mu_1,\mu_2,\dots)$ recording the number of boxes of weight $i = 1,2,\dots$ that the tableau contains, and the weight of a box is just the number that the box contains.
% For the sake of making a definition, we momentarily put some letters of our alphabet in unexpected roles, and ask you to suspend your disbelief. Given an arbitrary tableau $t$ having weight $w$ and shape $s$, we denote by $s(i)$ (resp.\ $w(i)$) the shape (resp.\ the weight) of $t(i)$, the tableau got from $t$ by discarding all boxes of weight exceeding $i$.

@@

The Lusztig datum $n_\bullet(\tau)$ of the tableau $\tau$ is a list of $ m(m-1)/2$ non-negative integers defined from its \textbf{Gelfand--Tsetlin pattern} $\gt(\tau) = (\lambda(i)_j)_{1\le j\le i\le m}$ by the formula 
$$
\begin{aligned}
n_\bullet(\tau)_{i,j} &= \lambda(i+1)_j - \lambda(i)_j = \text{ \# of boxes on the $j$th row of $\tau $ filled with $ i+1$}  \\
(i,j) &=  (1,1),(2,1),\dots,(m-1,1),%\\
(2,2),\dots,(m-1,2),%\\
\dots,%\\
(m-1,m-1)
\end{aligned}
% (i,j) = \begin{smallmatrix}
% \\
% (1,1) \\
% (2,1) & (2,2) \\
% \vdots & & \ddots \\
% (m-1,1) & (m-1,2) & \dots & (m-1,m-1) 
% \end{smallmatrix}
$$
% for 
% $(i,j) = (1,1),(2,1),\dots,(m-1,1),(2,2),\dots,(m-1,2),\dots,(m-1,m-1)$. 
% 
The pattern $\gt(\tau)$ is recorded as a lower-triangular matrix (the array of shapes of subtableaux $\tau(i)$) and the datum $n_\bullet(\tau)$ is recorded as a sequence. See \cite[\S 4]{berenstein1988tensor}. Below is an example.

@@

We can associate to $\tau$ an MV cycle as follows. Let    
%  
\begin{equation*}
    \mathring Z(\tau) = 
        \{
            L\in S^\mu : L_{\{1, \dots, i\}} \in\Gr^{\lambda(i)}\text{ for } i = 1, \dots m 
        \}
\end{equation*}
and $Z(\tau) = \overline{\mathring Z(\tau)}$. 
% 
\jcom{Maybe use $ [m] $ for $\{1, \dots, m\}$ ?}
\acom{And $\{1,\dots,i\}$ etc.}

The following result is closely related to Theorem 4 from \cite{dranowski2020generalized}.  It is also closely related to the description of MV cycles in terms of Kostant data obtained by Anderson--Kogan \cite{anderson2004mirkovic} (see Section 9 in \cite{kamnitzer2010mirkovic}).  We also note that a different map from Young tableaux to MV cycles was obtained by Gaussent--Littelmann--Nguyen \cite[Theorem 2]{gaussent2013knuth}.
\begin{proposition}
    \label{pr:newmvdes}
    $Z(\tau)$ is the MV cycle whose Lusztig datum with respect to $ \vi $ is $n_\bullet(\tau)$. 
\end{proposition}
\begin{proof}
    The proof of this theorem follows the same strategy as in \cite{dranowski2020generalized}.
\jcom{Maybe you would prefer to cite your thesis, Anne, but I think that it is better to cite the paper, since people can access it easier.}    
\acom{I've been meaning to update the paper. It's peppered with ``typos''.}

First, we consider \acom{add \texttt{mathring}'s to these $Z$'s?} \jcom{Sure.}
\begin{equation*}
Z(\tau)_1 := \{ L \in S^\mu \cap \Gr_+ : L_{\{1, \dots, i\}} \in S_+^{\lambda(i)} \text{ for } i = 1, \dots m\} \,. 
%Z(\tau)_2 := \{ L \in S^\mu : L \cap \cO^i \in S^{\lambda(i)} \cap \Gr^{\lambda(i)} \text{ for } i = 1, \dots m\}
\end{equation*}

Using Lemma \ref{le:Srdim}, we see that
$$
Z(\tau)_1 = \{ L \in \Gr_+ : \rdim L_{\{j, \dots, i\}} = \lambda(i)_j + \dots + \lambda(i)_i \text{ for all $ 1 \le j \le i \le m $} \}
$$

\jcom{We could write that $ \lambda(i)_j + \dots + \lambda(i)_i$ is the number of boxes on rows $ j, \dots, i$ filled with $ 1, \dots, i $.}
\acom{\emoji{thumbsup}}

From  the proof of \cite[Proposition 9.6]{kamnitzer2010mirkovic}, we see that $ Z(\tau)_1$ is equal to $ t^\mu A(n_\bullet(\tau))$, where $ A(n_\bullet)$ is defined in section 4.3 of \cite{kamnitzer2010mirkovic}.  Thus, $ \overline{Z(\tau_1)}$ is the MV cycle whose Lusztig datum is $ n_\bullet(\tau) $. (In \cite{kamnitzer2010mirkovic}, we worked with $  G(\cO) \backslash G(\cK)$, which we identify with $ \Gr $ using $ [g] \mapsto [g^{-1}]$.)

We also note that $Z(\tau_1) \cap \Gr^\lambda$ is dense in $ Z(\tau_1)$.  
\jcom{Need to justify this last sentence.  There are two possibilities.  Either we can use compute the MV polytope corresponds to $ \tau $ and show that it lies in the Weyl polytope for $ \lambda $ (is this done anywhere in the literature) or we could use the Mirkovic-Vybornov isomorphism.}

Fix $ i \in \{1, \dots, m\}$. If $ L \in S^\mu$, then $ L_{\{1, \dots, i\}} \in S^{\mu(i)}$.  Thus, we get a map $$ f_i : Z(\tau)_1 \rightarrow S^{\mu(i)} \cap S_+^{\lambda(i)} $$
From the definition of $ Z(\tau)_1 $, we see that $ f_i(Z(\tau_1) = Z(\tau(i)_1)$.  From above $ \Gr^{\lambda(i)} \cap f_i(Z(\tau(i)_1))$ is a dense constructible subset of $Z(\tau(i)_1)$.  Since $ Z(\tau)_1$ is irreducible, this implies that $ f_i^{-1}(S^{\mu(i)} \cap S_+^{\lambda(i)} \cap \Gr^{\lambda(i)}) $ is a dense constructible subset of $ Z(\tau_1) $.

Working with all $ i $ at once, we conclude that
$$ Z(\tau)_2 := \bigcap_{i=1}^m f_i^{-1}(S^{\mu(i)} \cap S_+^{\lambda(i)} \cap \Gr^{\lambda(i)}) $$
is a dense constructible subset of $ Z(\tau)_1$.  Thus $\overline{Z(\tau)_2} = \overline{Z(\tau)_1}$.

On the other hand, by definition 
$$
Z(\tau)_2 \subset \{ L \in S^\mu : L_{\{1, \dots, i\}} \in \Gr^{\lambda(i)} \text{ for } i = 1, \dots, m\}\,.
$$

@@

\jcom{Anne, can you reorganize this section to start with $ X(\tau) $ and then do the fusion afterwards?}

@@

Under the isomorphism of \Cref{cor:mvy} the MV cycle $Z(\tau)$ corresponds to the \new{generalized orbital varieties} $X(\tau)$ which is obtained by taking the closure of the top-dimensional 
% \acom{speak to this?} 
irreducible component of
% from $\tau\in YT(\lambda)_\mu$ by 
\[
\mathring X(\tau) = 
% \overline{
    \{
        A \in  \TT_\mu\cap\n : A\big|_{\CC^{|\mu(i)|}} \in \OO^{\lambda(i)} \text{ for each } i = 1,\dots,m
    \}\,. 
% }    
\]
% 
We do not know whether $\mathring X(\tau)$ is in general irreducible.
% reduced or has embedded components. 
\jcom{I guess we need to assume that $ \mu $ is dominant here.}

\begin{theorem}
    \label{pr:govsasirrecs}
    The irreducible components of $\overline{\OO}^\lambda\cap\TT_\mu\cap\n$ are $\{{X(\tau)} : \tau\in YT(\lambda)_\mu\}$. 
\end{theorem}

>>>>

%!TEX program = lualatex
% 
\documentclass{article} % [draft] TODO: use this at the end to check for overfills 
\usepackage{basic}
\usepackage{emoji}
% TODO: The matrix is lower triangular and the lower-triangular matrix is
% \setemojifont{Apple Color Emoji}
% Working title: 
\title{How to compute the fusion product of MV cycles in type $A$}
\author{Roger Bai, Anne Dranowski, Joel Kamnitzer} 
% \date{October 2020}
\date{\today}

@@

\begin{theorem} 
\label{th:OGrl}
There is an isomorphism
$$
    \overline{\OO}^{\lambda', \lambda''}_{0,\AA} \cap \TT_\mu \cong \overline\Gr^{\lambda', \lambda''}_{0,\AA} \cap \cW_\mu 
$$
given by $ (A,s) \mapsto (g(A)\CC[t]^m, s)$.
\end{theorem}
% \acom{Joel, you stopped writing $X_{0,\AA}$ in favour of $X_\AA$ --- should we do this everywhere?}
% \jcom{Oops, no I think that we should stick with $ X_{0, \AA}$.}
% TODO: comb the doc at the end for this notation 
\begin{proof}
Since we already have the isomorphism from \Cref{th:TmuWmu}, it suffices to show that for any $ A \in \TT_\mu$, 
$$ 
    (A,s) \in \overline{\OO}^{\lambda', \lambda''}_{0,\AA} \cap \TT_\mu \text{ if and only if } (g(A)\CC[t]^m, s) \in \overline\Gr^{\lambda', \lambda''}_{0,\AA} \cap \cW_\mu \,. 
$$
This follows immediately from \Cref{le:Grl1l2}.
\end{proof}
%
\acom{We need to assume that $\mu' +\mu''$ is dominant in the theorem below. The assumption for \Cref{th:TmuWmu} is also missing.} 
% 
\begin{theorem}
\label{th:OTGrW}
The isomorphism from \Cref{th:OGrl} 
% \acom{should ``Theorem'' etc.\ be capitalized when it appears in the middle of a sentence?} Yup.
restricts to an isomorphism
$$ 
    \overline{\OO}^{\lambda', \lambda''}_{0,\AA} \cap \UU^{\mu', \mu''}_{0,\AA} \cong \overline{\Gr}^{\lambda', \lambda''}_{0,\AA} \cap S^{\mu', \mu''}_{0,\AA}\,.
$$
\end{theorem}
% 
%\jcom{I was thinking about the proof of this result.  It is enough to prove that for $ A \in \OO^{\lambda', \lambda''}_s \cap \TT_\mu $, $g(A) \in S^{\mu', \mu''}_s $ if and only if $ A \in T^+$.  It is easy to show that if $ A \in \UU$, then $ g(A) \in S^{\mu', \mu''}_s $.  But the converse is not so obvious.  (I think Anne's paper/thesis might be incomplete on this point.)  I thought of two approaches: first to write down the lattice interpretation of $S^{\mu', \mu''}$ (similar to what is in my thesis in the Anderson-Kogan comparison section) or think about both $S^{\mu', \mu''}$ and $\UU$ as attracting sets for a $\CC^\times$ action.  Which do you prefer?}
% 
%\acom{Assuming Theorem 2 we think this follows by application of Roger's lemma. Also, I think my thesis is complete on this point.}
% 
%\jcom{I thought about this and I agree with you now.  I will write this up soon.} 
% 
%\acom{But I am also curious about the attracting set approach.}
% 
\begin{proof}
We could prove this by observing the both sides are the attracting locus of an appropriate $ \CC^\times$ action. However, we will give the following more algebraic proof.

@@

Restricting to the zero fibre and applying $\tau_0$ we recover \cite[Corollary~5.2.2]{dthesis}, itself a corollary of \cite[Theorem~4.4.1(c)]{mirkovic2019comparison}, the ordinary affine Grassmannian version of \Cref{th:OGrl}. 
% 
% \acom{The ordinary Gr should be called static. Really onomotopoeic. In opposition with moving. }
% 
\begin{corollary}
    \label{cor:mvy}
    $\overline{\OO}^\lambda \cap \TT_\mu\cap\n \cong\overline\Gr^\lambda\cap S^\mu_+$ where $\n\subset M_N(\CC)$ is the subalgebra of upper-triangular matrices. 
\end{corollary}
%The closures of the irreducible components of the righthand side are \textbf{MV cycles of type $\lambda$ and weight $\mu$}. By \cite{baumann2019mirkovic} the set of MV cycles of all possible types and weights can be stitched together to form a \textbf{perfect basis} of $\CC[N]$.

@@

Denote by $[a,b]$ the interval $\{a,a+1,\dots,b-1,b\}\subset\NN$. If $a = 1$ just write $[b]$, and if $a = b$ just write $\{a\}$. 
% If a = b? Just write $\{b\}$ 
We begin with the following definitions.
\begin{definition}
Let $ L \subset \cO^m$ be a point in $ \Gr_+$.  We define the relative dimension of $L $ by
$$
\rdim L := \dim_\CC \cO^m/L\,. 
$$
% 
If $ \gamma \subset [m]$, we define two lattices $ L^\gamma, L_\gamma $ in $ \cO^\gamma := \Sp_{\cO}(e_i : i \in \gamma) $ by
\begin{gather*}
%    \OO^\gamma := \Sp_{\OO}(e_i : i \in \gamma) \\
L^\gamma:= L \cap \cO^\gamma  \qquad L_\gamma := L / L^{\gamma^c} \subset \cO^m / \cO^{\gamma^c} = \cO^\gamma
\end{gather*}
where $ \gamma^c := [m] \setminus \gamma$.
% 
% \acom{I suggest replacing $\mathbb O$ by $R$, and $L/L^{\gamma^c}$ by $(L + R^{\gamma^c})/R^{\gamma^c}$ because that's what we're doing. In a lower triangular $g(A)$ columns ${i+1..m}$ cannot be used to eliminate entries in rows ${i+1..m}$ in the first $i$ columns over $\CC[t]$ or $\CC\xt$.}
% \jcom{If you think that it sounds better with $ \OO $ replaced by $ R$, then I won't object to the change. Note that the first sentence won't quite make sense ``let $ L \subset \OO^m$ be a point in $ \Gr_+$'', we would have to write ``let $ L \subset R^m$ be a lattice''.}
% \jcom{$ L /L^{\gamma^c} $ is the image of $ L $ in $ \OO^m / \OO^{\gamma^c}$ since the kernel of the map $ L \rightarrow \OO^m / \OO^{\gamma^c}$ is precisely $L^{\gamma^c} $, so I don't think that there is any problem here.  Again, if you think that it is better, I won't object to your suggested change.} 
% \jcom{Your question made me wonder about the following: how do we know that $ L_\gamma$ is a lattice?  (or $ L^\gamma$ for that matter)}
\end{definition}
We will also need the analogous definitions when $ \cO$ is replaced by $ \CC[t]$ and $L \in \Grth_+$.

We record the following easy observations.
\begin{lemma} \label{le:rdimTrans} Let $ L \in \Gr_+$.
\begin{enumerate}
    \item For any $ 
    \gamma \subset [m] $, we have $$ \rdim L_\gamma + \rdim L^{\gamma^c} = \rdim L \,.$$
    \item For any $ \gamma_1 \subset \gamma_2 \subset [m]$, we have $$ (L_{\gamma_1})_{\gamma_2} = L_{\gamma_2}
$$
\end{enumerate}
\end{lemma}

For any $ L \in \Gr_+ $ and $\gamma \subset [m]$, we let $ D_\gamma(L) := \rdim L^\gamma = \rdim L - \rdim L_{\gamma^c}$. By \cite[Proposition 9.3]{kamnitzer2010mirkovic}, this coincides with the general definition of $ D_\gamma$ given in that paper. \acom{We should introduce $D_\gamma$ briefly.}

Recall that $ S^\mu_- = N_-(\cK) t^\mu$. We also introduce the notation $S^\mu_+ := N(\cK)t^\mu$.
% 
% \jcom{Actually we will need this notation later when we talk about stable MV cycles.  Maybe we should change $ S^\mu $ to $S^\mu_-$.}
% \acom{For it.}

\begin{lemma} \label{le:Srdim}
    Let $ L \in \Gr_+$.  The following are equivalent:
    \begin{enumerate}
        \item $ L \in S_+^\mu$
        \item $ \rdim L^{[i]} = \mu_1 + \dots + \mu_i$ for all $ i = 1, \dots, m $.
        \item $ \rdim L_{[i+1,m]} = \mu_{i+1} + \dots + \mu_m$ for $ i  = 1, \dots, m$.
    \end{enumerate}
    Similarly, the following are equivalent:
        \begin{enumerate}
        \item $ L \in S_-^\mu$
        \item $ \rdim L^{[i+1,m]} = \mu_{i+1} + \dots + \mu_m$ for all $ i = 1, \dots, m $.
        \item $ \rdim L_{[i]} = \mu_{1} + \dots + \mu_i$ for $ i  = 1, \dots, m$.
    \end{enumerate}
\end{lemma}

We now introduce some notation related to Young tableaux. Denote by $YT(\lambda)_\mu$ the set of (possibly semistandard) Young tableaux of shape $\lambda$ and weight $\mu$. 
%Here, the weight of a tableau is a sequence $(\mu_1,\mu_2,\dots)$ recording the number of boxes of weight $i = 1,2,\dots$ that the tableau contains, and the weight of a box is just the number that the box contains.
% For the sake of making a definition, we momentarily put some letters of our alphabet in unexpected roles, and ask you to suspend your disbelief. Given an arbitrary tableau $t$ having weight $w$ and shape $s$, we denote by $s(i)$ (resp.\ $w(i)$) the shape (resp.\ the weight) of $t(i)$, the tableau got from $t$ by discarding all boxes of weight exceeding $i$.

@@

The Lusztig datum $n_\bullet(\tau)$ of the tableau $\tau$ is a list of $ m(m-1)/2$ non-negative integers defined from its \textbf{Gelfand--Tsetlin pattern} $\gt(\tau) = (\lambda(i)_j)_{1\le j\le i\le m}$ by the formula 
$$
\begin{aligned}
n_\bullet(\tau)_{i,j} &= \lambda(i+1)_j - \lambda(i)_j = \text{ \# of boxes on the $j$th row of $\tau $ filled with $ i+1$}  \\
(i,j) &=  (1,1),(2,1),\dots,(m-1,1),%\\
(2,2),\dots,(m-1,2),%\\
\dots,%\\
(m-1,m-1)
\end{aligned}
% (i,j) = \begin{smallmatrix}
% \\
% (1,1) \\
% (2,1) & (2,2) \\
% \vdots & & \ddots \\
% (m-1,1) & (m-1,2) & \dots & (m-1,m-1) 
% \end{smallmatrix}
$$
% for 
% $(i,j) = (1,1),(2,1),\dots,(m-1,1),(2,2),\dots,(m-1,2),\dots,(m-1,m-1)$. 
% 
The pattern $\gt(\tau)$ is recorded as a lower-triangular matrix (the array of shapes of subtableaux $\tau(i)$) and the datum $n_\bullet(\tau)$ is recorded as a sequence. See \cite[\S 4]{berenstein1988tensor}. Below is an example with $\lambda = (4,2)$ and $\mu = (3,2,1)$.

@@

We can associate to $\tau$ the MV cycle $Z(\tau)$ by taking the closure of the top-dimensional irreducible component of  
%  
\begin{equation*}
    \mathring Z(\tau) = 
        \{
            L\in S^\mu : L_{[i]} \in\Gr^{\lambda(i)}\text{ for } i = 1, \dots m 
        \}\,. 
\end{equation*}
% and $Z(\tau) = \overline{\mathring Z(\tau)}$. 
% 
% \jcom{Maybe use $ [m] $ for $\{1, \dots, m\}$ ?}
% \acom{And $\{1,\dots,i\}$ etc.}

The following result is closely related to Theorem 4 from \cite{dranowski2020generalized}.  It is also closely related to the description of MV cycles in terms of Kostant data obtained by Anderson--Kogan \cite{anderson2004mirkovic} (see Section 9 in \cite{kamnitzer2010mirkovic}).  We also note that a different map from Young tableaux to MV cycles was obtained by Gaussent--Littelmann--Nguyen \cite[Theorem 2]{gaussent2013knuth}. \acom{Many alsos}
\begin{proposition}
    \label{pr:newmvdes}
    $Z(\tau)$ is the MV cycle whose Lusztig datum (with respect to the standard reduced word) is $n_\bullet(\tau)$. 
\end{proposition}
\begin{proof}
    The proof of this theorem follows the same strategy as in \cite{dranowski2020generalized}.
% \jcom{Maybe you would prefer to cite your thesis, Anne, but I think that it is better to cite the paper, since people can access it easier.}    
% \acom{I've been meaning to update the paper. It's peppered with ``typos''.}

First, we consider 
% \acom{add \texttt{mathring}'s to these $Z$'s?} \jcom{Sure.}
\begin{equation*}
\mathring Z(\tau)_1 := \{ L \in S^\mu \cap \Gr_+ : L_{[i]} \in S_+^{\lambda(i)} \text{ for } i = 1, \dots m\} \,. 
%Z(\tau)_2 := \{ L \in S^\mu : L \cap \cO^i \in S^{\lambda(i)} \cap \Gr^{\lambda(i)} \text{ for } i = 1, \dots m\}
\end{equation*}

Using Lemma \ref{le:Srdim}, we see that
$$
\mathring Z(\tau)_1 = \{ L \in \Gr_+ : \rdim L_{[j,i]} = \lambda(i)_j + \dots + \lambda(i)_i \text{ for all $ 1 \le j \le i \le m $} \}
$$
where note that $ \lambda(i)_j + \dots + \lambda(i)_i$ is the number of boxes on rows $ j, \dots, i$ of weight $ 1, \dots, i $.
% \jcom{We could write that $ \lambda(i)_j + \dots + \lambda(i)_i$ is the number of boxes on rows $ j, \dots, i$ filled with $ 1, \dots, i $.}
% \acom{\emoji{thumbsup}}

From  the proof of \cite[Proposition 9.6]{kamnitzer2010mirkovic}, we see that $ \mathring Z(\tau)_1$ is equal to $ t^\mu A(n_\bullet(\tau))$, where $ A(n_\bullet)$ is defined in section 4.3 of \cite{kamnitzer2010mirkovic}. In particular it is irreducible. Thus, its closure is the MV cycle whose Lusztig datum is $ n_\bullet(\tau) $. (In \cite{kamnitzer2010mirkovic}, the third author worked with $  G(\cO) \backslash G(\cK)$, which we identify with $ \Gr $ using $ [g] \mapsto [g^{-1}]$.) \acom{Maybe we should avoid using square brackets as we haven't used them yet?}

We also note that $\mathring Z(\tau)_1\cap \Gr^\lambda$ is dense in $ \mathring Z(\tau)_1$.  
\jcom{Need to justify this last sentence.  There are two possibilities.  Either we can use compute the MV polytope corresponds to $ \tau $ and show that it lies in the Weyl polytope for $ \lambda $ (is this done anywhere in the literature) or we could use the Mirkovic-Vybornov isomorphism.}

Fix $ i \in \{1, \dots, m\}$. If $ L \in S^\mu_-$, then $ L_{[i]} \in S^{\mu(i)}_-$ so we get a map 
$$ 
f_i : \mathring Z(\tau)_1 \rightarrow S^{\mu(i)}_- \cap S_+^{\lambda(i)} \,. 
$$
From the definition of $ \mathring Z(\tau)_1 $, we see that $ f_i(\mathring Z(\tau)_1 )= \mathring Z(\tau(i))_1$.  
From above/again $ \Gr^{\lambda(i)} \cap f_i(\mathring Z(\tau(i))_1)$ is a dense constructible subset of $\mathring Z(\tau(i))_1$.  
Since $ \mathring Z(\tau)_1$ is irreducible, this implies that $ f_i^{-1}(S^{\mu(i)}_- \cap S_+^{\lambda(i)} \cap \Gr^{\lambda(i)}) $ \acom{$\cap \mathring Z(\tau)_1$?} is a dense constructible subset of $ \mathring Z(\tau)_1 $.

Working with all $ i $ at once, we conclude that
$$ 
\mathring Z(\tau)_2 := \bigcap_{i=1}^m f_i^{-1}(S^{\mu(i)}_- \cap S_+^{\lambda(i)} \cap \Gr^{\lambda(i)}) 
$$
is a dense constructible subset of $\mathring Z(\tau)_1$. \acom{Again are we maybe missing capping with $Z$ in the above $:=$?} Thus $\overline{\mathring Z(\tau)_2} = \overline{\mathring Z(\tau)_1}$.

On the other hand, by definition 
$$
\mathring Z(\tau)_2 \subset \mathring Z(\tau). % \{ L \in S^\mu : L_{[i]} \in \Gr^{\lambda(i)} \text{ for } i = 1, \dots, m\}\,.
$$

@@

% \jcom{Anne, can you reorganize this section to start with $ X(\tau) $ and then do the fusion afterwards?}

@@

Let $\tau\in YT(\lambda)_\mu$ with $\mu$ dominant. Under the isomorphism of \Cref{cor:mvy} the MV cycle $Z(\tau)$ corresponds to the \new{generalized orbital varieties} $X(\tau)$ which is obtained by taking the closure of the top-dimensional 
% \acom{speak to this?} 
irreducible component of
% from $\tau\in YT(\lambda)_\mu$ by 
\[
\mathring X(\tau) = 
% \overline{
    \{
        A \in  \TT_\mu\cap\n : A\big|_{\CC^{|\mu(i)|}} \in \OO^{\lambda(i)} \text{ for each } i = 1,\dots,m
    \}\,. 
% }    
\]
% 
We do not know whether $\mathring X(\tau)$ is in general irreducible. 
% reduced or has embedded components. 
% \jcom{I guess we need to assume that $ \mu $ is dominant here.}
\begin{lemma}
    Under the Mirkovi\'c--Vybornov isomorphism $\mathring X(\tau)$ is mapped isomorphically onto $\mathring Z(\tau)$. 
\end{lemma}
